### 2014-06-27 patchy-selection-paper.tex
changed back to my version of the decrease in rec
wrote chunk of discussion

\gc{Graham pasted this to the appendix as it felt like too much information to drop on people}
\plr{I am sympathetic, but it's only a paragraph\dots how about that last sentence, instad?}
%%%
% 
% \section{Limiting distribution of the transit time and hitchhiking haplotype}
% \label{apx:limiting_dist_of_HH_haplotype}
% 
% In the above we already assumed that the inter-patch distance $R$ is large relative to the migration-selection cline width $\sigma/\sqrt{s_m}$;
% also taking the large-$R$ limit, obtains a more intuitive description of the process
% via a central limit theorem for $\tau$.
% Define $\eta_\tau$ to be the deviation of $\tau$ from its mean, normalized by its standard deviation,
% so that $\tau = R/\sigma(2s_m)^{-1/2} + \eta_\tau \sqrt{R/\sigma} (2s_m)^{-3/4}$.
% Plugging this into equation~\eqref{eqn:haplen_cdf},
% after some algebra (and using that $\sqrt{R+\epsilon} = \sqrt{R} + \frac{\epsilon}{2\sqrt{R}} - \frac{\epsilon^2}{2\sqrt{R}^3} + O(\epsilon^3)$) we obtain that
% \begin{align}
%   \E[e^{-\ell \eta}] &= \exp\left\{ \frac{\ell^2}{2} + O\left(\frac{1}{R}\right) \right\}.
% \end{align}
% Since the Laplace transform uniquely determines a distribution \citep{durrett1996probability},
% this implies that $\eta_\tau$ converges in distribution to a standard Gaussian random variable as $R \to \infty$.
% In other words, we now know that the distribution of $\tau$ about its mean is approximately Gaussian.
% Since the standard deviation of $\tau$ is of lower order than its mean,
% a similar calculation shows that for large $R$, the haplotype halflength $L$ is approximately exponentially distributed, 
% with mean $\sqrt{2s_m} \sigma/R$.

@@

Based on this picture of
the lineage bouncing around in the patch,
we propose a simple heurstic to describe the probability of recombination. 
If recombination is on a slower time scale than mixing within the patch,
then the time until the last recombination is the integral of $r (1-p(x)) (1+s(x)/2)$ 
against the stationary distribution (equation~\ref{eqn:lineagestatdist}).
Within the patch, $p(x) \approx 1$,
so recombination off the $B$ background is unlikely to occur,
and so the rate of recombination depends to a first approximation 
on the proportion of the patch in which $p(x)$ is intermediate,
i.e.\ where both $B$ and $b$ are present.
This region where $p(x)$ is intermediate has width of order $\sigma/\sqrt{s}$, where $s = \min(s_p,s_m)$.
As we increase the size of a patch, the area on which $p(x)$ is intermediate scales with the perimeter of the patch.
Therefore, the proportion of the time a lineage at stationarity is found in the intermediate regions
multiplied by $p(x)$ in these intermediate regions, where we assume
$p(x) \approx 1/2$. Under these approximations the per generation rate of
recombination of a linked lineage to the b background is
reduced by a factor roughly equal to the proportion of the patch lying within distance $\sigma/\sqrt{s}$ of the boundary,
i.e.
% \begin{equation}
% r (\sigma/\sqrt{s})A^{-1/d} \label{eqn:rate-rec-in-patch}
% \end{equation}
\plr{Note to ourselves: If a $d$-dimensional object with area A has radius $r$ (in any sense)
and you shrink it to have radius $r_0$,
then the resulting object has area $A \times (r_0/r)^d$;
here we have $r_0 = r - \sigma/\sqrt{s}$,
in $d=1$ the radius is $A/2$ and in $d=2$ it is $\sqrt{A/\pi}$.}
\begin{align}
  r_\text{eff} &=
    \begin{cases}
      r \frac{2\sigma}{A\sqrt{s}}    \qquad & d=1 \\
      r \frac{\sqrt{\pi} \sigma }{ \sqrt{ s A } } \qquad & d=2 \\  
\end{cases} ,
\label{eqn:rate-rec-in-patch}   %(\sigma/\sqrt{s})A^{-1/d}
\end{align}
i.e.\ the reduction in recombination rate is proportional to $(\sigma/\sqrt{s})A^{-1/d}$.
We can use this to think about the length of the haplotype
associated with the $B$ allele after a given length of time $T$,
for instance, the time since the $B$ allele joined the patch.
\plr{changed this to genetic lengths so as not to introduce another parameter, $r_{BP}$}
% If the recombination rate per base is $r_{BP}$, 
% then the length of the haplotype associated with the $B$ allele is
% \begin{equation}
% \frac{\sqrt{s}}{\sigma A^{-1/d}}\frac{1}{r_{BP}}
% \end{equation} 
In units of genetic length (Morgans),
the length of the haplotype associated with the $B$ allele is, up to a constant,
\begin{equation}
  \frac{A^{1/d}\sqrt{s}}{\sigma T}
\end{equation} 
Note that the length of haplotype decays at a rate proportional to the
cline width $\sqrt{s} / \sigma$, as the narrower the cline the less
opportunity for recombination to move our lineage off the $B$ background.

@@

>>>>

% 
% \section{Limiting distribution of the transit time and hitchhiking haplotype}
% \label{apx:limiting_dist_of_HH_haplotype}
% 
% In the above we already assumed that the inter-patch distance $R$ is large relative to the migration-selection cline width $\sigma/\sqrt{s_m}$;
% also taking the large-$R$ limit, obtains a more intuitive description of the process
% via a central limit theorem for $\tau$.
% Define $\eta_\tau$ to be the deviation of $\tau$ from its mean, normalized by its standard deviation,
% so that $\tau = R/\sigma(2s_m)^{-1/2} + \eta_\tau \sqrt{R/\sigma} (2s_m)^{-3/4}$.
% Plugging this into equation~\eqref{eqn:haplen_cdf},
% after some algebra (and using that $\sqrt{R+\epsilon} = \sqrt{R} + \frac{\epsilon}{2\sqrt{R}} - \frac{\epsilon^2}{2\sqrt{R}^3} + O(\epsilon^3)$) we obtain that
% \begin{align}
%   \E[e^{-\ell \eta}] &= \exp\left\{ \frac{\ell^2}{2} + O\left(\frac{1}{R}\right) \right\}.
% \end{align}
% Since the Laplace transform uniquely determines a distribution \citep{durrett1996probability},
% this implies that $\eta_\tau$ converges in distribution to a standard Gaussian random variable as $R \to \infty$.
% In other words, we now know that the distribution of $\tau$ about its mean is approximately Gaussian.
% Since the standard deviation of $\tau$ is of lower order than its mean,
% a similar calculation shows that for large $R$, the haplotype halflength $L$ is approximately exponentially distributed, 
% with mean $\sqrt{2s_m} \sigma/R$.

@@

Based on this picture of
the lineage bouncing around in the patch,
we propose a simple heurstic to describe the probability of recombination. 
If recombination is on a slower time scale than mixing within the patch,
then the time until the last recombination is the integral of $r (1-p(x)) (1+s(x)/2)$ 
against the stationary distribution (equation~\ref{eqn:lineagestatdist}).
Within the patch, $p(x) \approx 1$,
so recombination off the $B$ background is unlikely to occur,
and so the rate of recombination depends to a first approximation 
on the proportion of the patch in which $p(x)$ is intermediate,
i.e.\ where both $B$ and $b$ are present.
This region where $p(x)$ is intermediate has width of order $\sigma/\sqrt{s}$, where $s = \min(s_p,s_m)$.
As we increase the size of a patch, the area on which $p(x)$ is intermediate scales with the perimeter of the patch.
Therefore, the proportion of the time a lineage at stationarity is found in the intermediate regions
multiplied by $p(x)$ in these intermediate regions, where we assume
$p(x) \approx 1/2$. Under these approximations the per generation rate of
recombination of a linked lineage to the b background is
reduced by a factor roughly equal to the proportion of the patch lying within distance $\sigma/\sqrt{s}$ of the boundary,
i.e.
% \begin{equation}
% r (\sigma/\sqrt{s})A^{-1/d} \label{eqn:rate-rec-in-patch}
% \end{equation}
\plr{Note to ourselves: If a $d$-dimensional object with area A has radius $r$ (in any sense)
and you shrink it to have radius $r_0$,
then the resulting object has area $A \times (r_0/r)^d$;
here we have $r_0 = r - \sigma/\sqrt{s}$,
in $d=1$ the radius is $A/2$ and in $d=2$ it is $\sqrt{A/\pi}$.}
\begin{align}
r_{eff} &=
    \begin{cases}
    r \frac{ \sigma /\sqrt{s} }{ \sigma/\sqrt{s} + A }    \qquad & d=1 \\
   r \frac{\pi  \sigma /\sqrt{s}  \left( \sigma /\sqrt{s}+
  \sqrt{A/\pi} \right)}{A+\pi \sigma/\sqrt{s} \left( \sigma/\sqrt{s} +
  \sqrt{A/\pi} \right) }    \qquad & d=2 \\  
\end{cases}
\label{eqn:rate-rec-in-patch}   %(\sigma/\sqrt{s})A^{-1/d}
\end{align}
\gc{Altered the above.}
which scales the recombination rate roughly with $(\sigma/\sqrt{s})A^{-1/d}$.
We can use this to think about the length of the haplotype
associated with the $B$ allele after a given length of time $T$,
for instance, the time since the $B$ allele joined the patch.
\plr{changed this to genetic lengths so as not to introduce another parameter, $r_{BP}$}
% If the recombination rate per base is $r_{BP}$, 
% then the length of the haplotype associated with the $B$ allele is
% \begin{equation}
% \frac{\sqrt{s}}{\sigma A^{-1/d}}\frac{1}{r_{BP}}
% \end{equation} 
In units of genetic length (Morgans),
the length of the haplotype associated with the $B$ allele is, up to a constant,
\begin{equation}
  \frac{A^{1/d}\sqrt{s}}{\sigma T}
\end{equation} 
Note that the length of haplotype decays at a rate proportional to the
cline width $\sqrt{s} / \sigma$, as the narrower the cline the less
opportunity for recombination to move our lineage off the $B$ background.

@@

We have developed a simple set of approximations to describe the rate
at which organisms in a patchy environment adapt through novel
mutations or migration from other nearby patches (eqn.
\eqref{eqn:mutrate} and \eqref{eqn:migrate}), and the relative
probability of each of these outcomes (eqn. \eqref{eqn:parallel_prob}). 
We have also determined the characteristic genomic length scales over which
the various signals of hitchhiking should be visable
(eqn. \eqref{eqn:haplotype_length} and ).

These results show that the probability of parallel adaptation is
independent of both the population density and the beneficial
selection coefficient. This follows from the fact that these
parameters both increase the rate of adaptation via both migration and
mutation equally. Thus they are important for the time-scale over
which adaptation occurs but not the mode of adaptation. This
constrasts with models of convergent adaptation within panmictic
populations \citep{} and geographically spread populations adapting to
homogeneous populations \citep{RalphCoop}, where the number of
adaptive alleles increases with the population size/density. In all of these models the dependence on the beneficial selection
coefficient is absent or weak, due to the fact that selection both
aids establiishment of new alleles and the spread of existing
alleles.

The primary determinant of the probability of parallel adaptation is
$\sigma/\sqrt{s_d}$, the characteristic geographic distance of the
allele frequency decay off the patch. We find that patches that are close together on
this scale will adapt via migration, but there will be a sharp
transition to convergent adaptation as we move to patches that are further apart.

In coming to this conclusion, we have ignored the possibility of very long distance migration,
instead focusing local dispersal (as parameterized by a Gaussian
dispersal model). However, dispersal distributions can be very heavy
tailed, with a small fraction of individuals moving very long distances indeed, in
addition over long time-scales very rare chance events (mice sucked up in
hurricanes etc) could play a role in spreading migrant alleles. Such
tail events could greatly increase the probability of parallel
adaptation above that predicted by our model. Furthermore, if
adaptive alleles do move between distant patches via rare long distance
migration they will be associated with a much longer shared haplotype than
predicted by eqn. \eqref{eqn:haplotype_length}. As such we view our
results as a null model by which the contribution of long distribution
migrants to adaptation could be empirically judged.

We have also assumed that the alleles have a fitness effect in
heterozygotes both on and off the patch. However, 
 recessive alleles could also contribute to adaptation to patchy
 environments. \gc{should we cite Nachman eg here}.  Our
 simple branching process framework does not easily accomodate recessive
 alleles as it ignores the selection interaction of alleles, and so a
 different approach will be needed. At first sight it
 seems like recessive alleles may be able to contribute to adaptation
 via migration over longer distances, as the migrant alleles will be at low
 frequency between patches and so usually invisible to
 selection. However, this fact also means that they are unlikely to
 establish on the beneficial patch upon reaching it via migration. As
 such it is not clear what our intuition should be about the
 contribution of recessive alleles to adaptation via migration.

\gc{will write a couple more broader paragraphs}

The falling cost of sequencing over the next decade offers the
opportunity to investigate the genomics of parallel adaptation over a
large s
Over the coming decade population genomic data will be generated
